---
output:
  html_document: default
  pdf_document: default
---

|                                    |
|------------------------------------|
| Title: "NEE gap-filling and partitioning" |
| Author: "Fabio Turco"              |
| Date: "2024-04-24"                 |
| Output:                            |
| pdf_document: default              |
| html_document: default             |

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

**Script to detect seasonal Ustar thresholds from level-3.2 fluxes.\
**In addition, the script is doing a preliminary MDS gap-filling of NEE which won't be used in the next steps as machine learning approaches have been shown to be less biased.

HIGHEST QUALITY ONLY (QCF=0) FLUXES (\_L3.1_L3.2_QCF0), quality-controlled with Level-2 and Level-3.2 flags, including Level-3.1 storage correction are used for the ustar detection.

## LIBRARIES
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(REddyProc)
library(caTools)
library(data.table)
library(viridis)
```

Define the run ID
```{r}
run_id <- format(Sys.time(), "%Y%m%d",tz="GMT")
print(run_id)
```

## SOURCE FILE

```{r}
file_fluxes_meteo <- "../../80_GAP-FILLING/81.2_GapFillingSubsetForREddyProc.csv"
```

## DATA FROM FILE

-   Timestamp is in Etc/GMT-1
-   Need to provide an end timestamp because EddyProc is shifting to the middle internally by subtracting 15 minutes

```{r}
filedata <- read_csv(file_fluxes_meteo, show_col_types = FALSE)
filedata$TIMESTAMP_END <- as.POSIXct(filedata$TIMESTAMP_END, tz = "Etc/GMT-1")
```
## TIME INTERVAL
```{r}
filedata <- filedata %>% filter(TIMESTAMP_END >= as.POSIXct('2023-11-07 17:00:00', tz = 'Etc/GMT-1')
                                & TIMESTAMP_END <= as.POSIXct('2025-06-05 01:00:00', tz = 'Etc/GMT-1'))

summary(filedata[c('TIMESTAMP_END', 'NEE_L3.3_CUT_50_QCF', 'sw_in', 'ta', 'vpd', 'parcel', 'parcel_certainty', 'SW_IN_POT')], 3)
```
## DATA COLUMNS

**Using QCF fluxes**
```{r}
make_parcel_df <- function(letter, pot_thresh = 20) {
  filedata %>%
    mutate(night = !is.na(SW_IN_POT) & SW_IN_POT <= pot_thresh) %>%   # night definition
    transmute(
      TIMESTAMP = TIMESTAMP_END,
      NEE  = case_when(
        night ~ NEE_L3.3_CUT_50_QCF,                                   # keep at night for all
        parcel == letter & parcel_certainty == "certain" ~ NEE_L3.3_CUT_50_QCF,
        TRUE ~ NA_real_
      ),
      Rg   = sw_in,
      Tair = ta,
      VPD  = vpd
    )
}

EddyData_A <- make_parcel_df("A")       # night kept + A certain
EddyData_B <- make_parcel_df("B")       # night kept + B certain
```

## EDDYPROC CONFIGURATION

### Initialize R5 reference class
```{r}
EddyProc_A <- sEddyProc$new('CH-TAN_A',  EddyData_A, c('NEE', 'Rg','Tair', 'VPD'), ColPOSIXTime = "TIMESTAMP")
EddyProc_A$sSetLocationInfo(LatDeg = 47.480620, LongDeg = 8.911868, TimeZoneHour = 1)

EddyProc_B <-sEddyProc$new('CH-TAN_B',  EddyData_B, c('NEE', 'Rg','Tair', 'VPD'), ColPOSIXTime = "TIMESTAMP")
EddyProc_B$sSetLocationInfo(LatDeg = 47.480620, LongDeg = 8.911868, TimeZoneHour = 1)

head(EddyProc_A$sDATA)
head(EddyProc_B$sDATA)
```
## MDS GAP-FILLING

### NEE

**I have not dealt with the issue of negative values filled by MDS during night**
```{r}
#+++ Gap fill variable with default variables and limits
EddyProc_A$sMDSGapFill("NEE")
EddyProc_B$sMDSGapFill("NEE")
```
```{r}
EddyProc_A$sPlotHHFluxes("NEE_f")
png("plots/NEE_MDS_Fingerprint_parcelA.png", width = 900, height = 700, res = 120)
EddyProc_A$sPlotFingerprintY("NEE_f", Year = 2024)
dev.off()
EddyProc_B$sPlotHHFluxes("NEE_f")
png("plots/NEE_MDS_Fingerprint_parcelB.png", width = 900, height = 700, res = 120)
EddyProc_B$sPlotFingerprintY("NEE_f", Year = 2024)
dev.off()
```

```{r}
# helper: convert REddyProc export -> tidy frame (keeps your timestamp)
as_plotdf <- function(out, label) {
  tibble(
    DateTime = filedata$TIMESTAMP_END,
    NEE_orig = out$NEE_orig,
    NEE_f    = out$NEE_f,
    parcel   = label
  ) %>%
    filter(!is.na(NEE_f))
}

# build combined data
outA <- EddyProc_A$sExportResults()
outB <- EddyProc_B$sExportResults()
plotdf <- bind_rows(
  as_plotdf(outA, "Parcel A"),
  as_plotdf(outB, "Parcel B")
)

# shared y-limits across both panels
ylims <- range(c(plotdf$NEE_f, plotdf$NEE_orig), na.rm = TRUE)

# plot: grey filled line + black measured points; side-by-side with shared y
ggplot(plotdf, aes(DateTime)) +
  geom_line(aes(y = NEE_f), color = "grey50", linewidth = 0.4, alpha = 0.6) +
  geom_point(data = subset(plotdf, !is.na(NEE_orig)),
             aes(y = NEE_orig), pch = 16, size = 0.9, alpha = 0.95) +
  facet_grid(. ~ parcel, scales = "free_x") +   # side-by-side, same y by default
  coord_cartesian(ylim = ylims) +               # enforce identical y-range
  scale_x_datetime(date_breaks = "3 month", date_labels = "%Y-%m") +
  labs(x = NULL, y = expression(NEE~(mu*mol~m^{-2}~s^{-1}))) +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))
```
## NEE PARTITIONING

```{r}
# We first need to gap-fill Tair, otherwise it could give error, although it should be without gaps
EddyProc_A$sMDSGapFill('Tair', FillAll = FALSE)
EddyProc_B$sMDSGapFill('Tair', FillAll = FALSE)
# NIGHT-TIME PARTITIONING
EddyProc_A$sMRFluxPartition()
EddyProc_A$sPlotFingerprint('GPP_f')
EddyProc_B$sMRFluxPartition()
EddyProc_B$sPlotFingerprint('GPP_f')
```
```{r}
outA <- EddyProc_A$sExportResults()
outB <- EddyProc_B$sExportResults()

plot_gpp <- bind_rows(
  tibble(DateTime = filedata$TIMESTAMP_END, GPP_f = outA$GPP_f, parcel = "Parcel A"),
  tibble(DateTime = filedata$TIMESTAMP_END, GPP_f = outB$GPP_f, parcel = "Parcel B")
) %>%
  filter(!is.na(GPP_f))

ylims <- range(plot_gpp$GPP_f, na.rm = TRUE)

ggplot(plot_gpp, aes(DateTime, GPP_f)) +
  geom_line(linewidth = 0.4) +
  facet_grid(. ~ parcel, scales = "free_x") +  # same y by default
  coord_cartesian(ylim = ylims) +
  scale_x_datetime(date_breaks = "4 month", date_labels = "%Y-%m") +
  labs(x = NULL, y = "GPP_f") +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))
```


